---
title: "Descriptive analysis"
author: "Yuan Liao"
date: "10/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs, include=TRUE, warning=FALSE, message=FALSE}
library(tm)
library(qdap)

stats <- list()
stats$lengths <- list()
stats$lines <- list()

glimpse <- function(var, stats){
  f <- paste0("../dbs/en_US/en_US.", var, ".txt")
  con <- file(f, "r")
  # Take the first three lines to take a look
  l <- readLines(con=con, 3)
  
  # Count how many lines in total
  len <- length(readLines(con=con, encoding = 'UTF-8'))
  close(con=con)
  
  # Find the longest lines
  con <- file(f, "r")
  L <- lapply(readLines(con=con, encoding = 'UTF-8'), nchar)
  L.max <- max(unlist(L), na.rm = TRUE)
  close(con=con)
  
  # Close the file connection and print the first three lines
  stats$lengths[[var]] <- len
  stats$lines[[var]] <- L.max
  print(l)
  return(stats)
}
```
This document takes dbs/en_US/ as an example to explore data format and compute the basic statistics of the dataset.

## 1 Data format
There are three sources of the data: blogs, news, and twitter.

### 1-1 en_US.blogs.txt
```{r data blog-samples, warning=FALSE, message=FALSE}
stats <- glimpse('blogs', stats)

```

### 1-2 en_US.news.txt
```{r data news-samples, warning=FALSE, message=FALSE}
stats <- glimpse('news', stats)
```

### 1-3 en_US.twitter.txt
```{r data twitter-samples, warning=FALSE, message=FALSE}
stats <- glimpse('twitter', stats)
```

### 2 Statistics

|         | No. of lines              | No. of characters of the longest line |
|---------|---------------------------|---------------------------------------|
| blogs   | `r stats$lengths$blogs`   | `r stats$lines$blogs`                 |
| news    | `r stats$lengths$news`    | `r stats$lines$news`                  |
| twitter | `r stats$lengths$twitter` | `r stats$lines$twitter`               |

### 3 Preliminary word/sentence matching in en_US.twitter.txt
#### 3-1 Occurance of "love" and "hate"
```{r tweets-love-hate}
var <- 'twitter'
f <- paste0("../dbs/en_US/en_US.", var, ".txt")

# Find the occurance of a certain word
word_match <- function(f){
  con <- file(f, "r")
  l <- unlist(lapply(readLines(con=con, encoding = 'UTF-8'), function(x){grepl("love", x)}))
  word.count <- length(l[l==TRUE])
  close(con=con)
  return(word.count)
}


```
